---
title: "P8105 HW2"
author: "Hansheng Zhang"
date: 2021-10-07
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

# Problem 1

Mr. Trash Wheel - Read and clean Mr. Trash Wheel data set

``` {r}
trashwheel_df = read_excel("./Data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N535") %>% # omit header
  janitor::clean_names() %>% 
  drop_na("dumpster") %>% 
  mutate(sports_balls = as.integer( #convert to integer vector
    round(sports_balls, digits = 0)))
```
The following code imports the excel sheets named "2018 Precipitation" and "2019 Precipitation". 

```{r}
precip_2018 =
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2018 Precipitation", range = "A2:B15") %>%
  janitor::clean_names() %>% 
  mutate(year = "2018")

precip_2019 =
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2019 Precipitation", range = "A2:B15") %>%
  janitor::clean_names() %>% 
  mutate(year = "2019")
```

Combining `precip_2018` and `precip_2019` tibbles into one and mutating the `month` vector from a `dbl` to a `chr`.

```{r}
precip_2018_2019 =
    bind_rows(precip_2018, precip_2019) %>% 
    mutate(month = month.name[month])
```

After reading and cleaning the Mr. Trash Wheel data set, the total number of observations is now `r nrow(trashwheel_df)`. There are `r ncol(trashwheel_df)` variables in the cleaned Mr. Trash Wheel data. Key variables in the Mr. Trash Wheel data are dumpster, month, year, and sports balls. 

The number of observations in the combined precipitation data set is `r nrow(precip_2018_2019)`. There are `r ncol(precip_2018_2019)` variables in the combined precipitation data set, which are `r names(precip_2018_2019)`.

The total precipitation in the year 2018 was `r sum(pull(filter(precip_2018_2019, year == 2018), total))`.

The median number of sports balls in a dumpster in 2019 was `r median(pull(filter(trashwheel_df, year == 2019), sports_balls))`. 

# Problem 2

538.com - Read and merge data sets

```{r}
## Import 3 csv files we are using

pols_month = read_csv("./data/pols-month.csv")
snp = read_csv("./data/snp.csv")
unemployment = read_csv("./data/unemployment.csv")
```

```{r}
## Clean pols_month by separating mon into 3 character variables and creating president varaible to take on dem or gop. 

pols_month =
  pols_month %>% 
  separate(mon, into = c("year", "month", "day"), convert = TRUE) %>%
  mutate(
    month = month.name[month],
    president = recode(prez_dem, `1` = "dem", `0` = "gop")) %>% 
  select(-prez_gop, -prez_dem, -day)
```

```{r}
## Clean snp in a similar manner and rearrange columns

snp =
  snp %>%
  separate(date, into = (c("month", "day", "year")), convert = TRUE) %>%
  mutate(month = month.name[month]) %>%
  relocate(year, month) %>% 
  select(-day)
```

```{r}
## Tidying unemployment data to be merged with previous two data sets. Switching from wide to long formats

unemployment =
  unemployment %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment") %>%
  mutate(    ## Converting to same format
    Year = as.integer(Year),
    Month = match(month, month.abb),
    Month = month.name[Month]) %>% 
  janitor::clean_names()
```

```{r}
## Merging data sets by first joining pols_month with snp, and then joining unemployment with that result

pols_snp = left_join(pols_month, snp, by = c("year", "month"))

pols_snp_unemployment = left_join(pols_snp, unemployment, by = c("year","month"))
```

The data sets `pols_months`, `snp`, and `unemployment` contained US political office, stock market, and unemployment data, respectively. After cleaning, the `pols_month` data set contained `r ncol(pols_month)` variables, key ones including year, month, president, close, and unemployment. The year range for the data covered from `r min(pull(pols_snp_unemployment, year))` through `r max(pull(pols_snp_unemployment, year))`.


# Problem 3

Baby names - Tidying popular baby names

```{r}
baby_names =
  read_csv("./data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(
    childs_first_name = str_to_lower(childs_first_name),
    gender = str_to_lower(gender),
    gender = as.factor(gender),
    ethnicity = str_to_lower(ethnicity),
    ethnicity = as.factor(ethnicity),
    ethnicity = recode(ethnicity,
                       "asian and paci" = "asian and pacific islander",
                       "black non hisp" = "black non hispanic",
                       "white non hisp" = "white non hispanic"),
    year_of_birth = as.integer(year_of_birth)) %>% 
  distinct()
```

The following code generates a table showing the ranking of the name `Olivia` among each ethnicity for the years 2011 through 2016.

```{r}
baby_olivia <-
  filter(baby_names, childs_first_name == "olivia") %>%
  select(-c("gender", "childs_first_name", "count")) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank") %>% 
  knitr::kable()

baby_data_olivia
```

The following code generates a table showing the most popular `male` names among each ethnicity for the years 2011 through 2016.

```{r generate_table_male_names}
baby_data_males <-
  filter(baby_data,
    rank == 1,
    gender == "male") %>%
  select(-c("gender", "count", "rank")) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name") %>% 
  knitr::kable(align = 'c')

baby_data_males
```

The following code generates a scatterplot showing the number of children with a name against the rank in popularity of that name, after filtering for `white non hispanic`, `male`, and `2016` as the `year_of_birth`. 

```{r scatterplot_of_names, fig.align= 'center'}
baby_plot <-
    filter(baby_data,
           ethnicity == "white non hispanic",
           gender == "male",
           year_of_birth == 2016)
bdp <- ggplot(baby_data_plot, aes(x = rank, y = count)) + geom_point(colour = "red", size = 2, alpha = 1/5)
bdp + labs(
    title = "The Total Count of Particular Names* by Their Popularity Rank",
    subtitle = "Amongst White Non-Hispanic Males Born in 2016",
    x = "Popularity Ranking",
    y = "Number of Male Babies Named a Particular Name",
    caption = "*names not shown")
```